---
title: "DA_401_Data_Exploration-Confounders"
author: "John Platt"
date: "2025-10-13"
output: 
  html_document:
    code_folding: hide
---

## Introduction
We perform the remainder of the third goal of our data exploration on the 2021-23 NSDUH data, using the following methods: weighted chi-squared tests and weighted 2 sample t-tests. As a reminder, the third goal of our data exploration is to see the association between confounders and suicidal ideation.

## Weighted 2 Sample t-tests
We see the association between our numerical confounders and suicidal ideation

```{r specifying our survey design, message=FALSE, warning=FALSE}
#Load in cleaned dataset from data folder of repository
nsduh_2123<-read.csv("C:/Users/John Platt/OneDrive/DA301Project/John-Platt-DA-401-Project/data/NSDUH_2021_2023.csv")
#Install survey package
library(survey) #Library for survey design

options(survey.lonely.psu = "adjust")  #common & documented choice to adjust variance if there are single psus within stratum

#Specify survey design
des <- svydesign(
  id      = ~VEREP, #PSU/cluster id
  strata  = ~VESTR_C, #variance strata
  weights = ~ANALWT2_C3, #analysis weight
  data    = nsduh_2123, #2021-2023 NSDUH dataframe
  nest = TRUE
)
```

```{r weighted 2 sample t-test between alcohol use and suicidal thoughts for young adults, include=TRUE}
#Compare means within CATAGE==2
tt <- svyttest(IRALCFY ~ IRSUICTHNK, design = subset(des, CATAGE == 2))
tt
svyby(
  ~IRALCFY, #outcome variable
  ~IRSUICTHNK, #grouping variable (two categories)
  subset(des, CATAGE == 2), #subset population
  svymean, #survey weighted mean
  vartype = c("se", "ci"), #include SEs and 95% CIs
  na.rm = TRUE 
)
#Extract overall standard deviation of IRALCFY
svy_var  <- svyvar(~IRALCFY, design = subset(des, CATAGE == 2))
# Extract standard deviation
svy_sd <- sqrt(as.numeric(svy_var))
svy_sd
#Extract standard deviation by group
svy_byvar <- svyby(~IRALCFY, ~IRSUICTHNK, design = subset(des, CATAGE == 2), FUN = svyvar, na.rm = TRUE)
svy_byvar$sd <- sqrt(svy_byvar$IRALCFY)
svy_byvar
##Survey-weighted Mann–Whitney (Wilcoxon rank-sum) test
svyranktest(IRALCFY ~ IRSUICTHNK, design = subset(des, CATAGE == 2), test = "wilcoxon")
```

```{r weighted 2 sample t-test between cigarette use and suicidal thoughts for young adults, include=TRUE}
#Compare means within CATAGE==2
tt <- svyttest(IRCIGFM ~ IRSUICTHNK, design = subset(des, CATAGE == 2))
tt
svyby(
  ~IRCIGFM, #outcome variable
  ~IRSUICTHNK, #grouping variable (two categories)
  subset(des, CATAGE == 2), #subset population
  svymean, #survey weighted mean
  vartype = c("se", "ci"),   #include SEs and 95% CIs
 na.rm = TRUE 
  )
#Extract overall standard deviation of IRCIGFM
svy_var  <- svyvar(~IRCIGFM, design = subset(des, CATAGE == 2))
# Extract standard deviation
svy_sd <- sqrt(as.numeric(svy_var))
svy_sd
#Extract standard deviation by group
svy_byvar <- svyby(~IRCIGFM, ~IRSUICTHNK, design = subset(des, CATAGE == 2), FUN = svyvar, na.rm = TRUE)
svy_byvar$sd <- sqrt(svy_byvar$IRCIGFM)
svy_byvar
##Survey-weighted Mann–Whitney (Wilcoxon rank-sum) test
svyranktest(IRCIGFM ~ IRSUICTHNK, design = subset(des, CATAGE == 2), test = "wilcoxon")
```

```{r weighted 2 sample t-test between binge drinking and suicidal thoughts for young adults, include=TRUE}
#Compare means within CATAGE==2
tt <- svyttest(IRALCBNG30D ~ IRSUICTHNK, design = subset(des, CATAGE == 2))
tt
svyby(
  ~IRALCBNG30D, #outcome variable
  ~IRSUICTHNK, #grouping variable (two categories)
  subset(des, CATAGE == 2), #subset population
  svymean, #survey weighted mean
  vartype = c("se", "ci"),   #include SEs and 95% CIs
  na.rm = TRUE  
  )
#Extract overall standard deviation of IRALCBNG30D
svy_var  <- svyvar(~IRALCBNG30D, design = subset(des, CATAGE == 2))
# Extract standard deviation
svy_sd <- sqrt(as.numeric(svy_var))
svy_sd
#Extract standard deviation by group
svy_byvar <- svyby(~IRALCBNG30D, ~IRSUICTHNK, design = subset(des, CATAGE == 2), FUN = svyvar, na.rm = TRUE)
svy_byvar$sd <- sqrt(svy_byvar$IRALCBNG30D)
svy_byvar
##Survey-weighted Mann–Whitney (Wilcoxon rank-sum) test
svyranktest(IRALCBNG30D ~ IRSUICTHNK, design = subset(des, CATAGE == 2), test = "wilcoxon")
```

```{r weighted 2 sample t-test between nicotine vaping and suicidal thoughts for young adults, include=TRUE}
#Compare means within CATAGE==2
tt <- svyttest(IRNICVAP30N ~ IRSUICTHNK, design = subset(des, CATAGE == 2 & IRNICVAP30N != -9))
tt
svyby(
  ~IRNICVAP30N, #outcome variable
  ~IRSUICTHNK, #grouping variable (two categories)
  subset(des, CATAGE == 2 & IRNICVAP30N != -9), #subset population
  svymean, #survey weighted mean
  vartype = c("se", "ci"),   #include SEs and 95% CIs
na.rm = TRUE 
  )
#Extract overall standard deviation of IRNICVAP30N
svy_var  <- svyvar(~IRNICVAP30N, design = subset(des, CATAGE == 2), na.rm = TRUE)
# Extract standard deviation
svy_sd <- sqrt(as.numeric(svy_var))
svy_sd
#Extract standard deviation by group
svy_byvar <- svyby(~IRNICVAP30N, ~IRSUICTHNK, design = subset(des, CATAGE == 2), FUN = svyvar, na.rm = TRUE)
svy_byvar$sd <- sqrt(svy_byvar$IRNICVAP30N)
svy_byvar
##Survey-weighted Mann–Whitney (Wilcoxon rank-sum) test
svyranktest(IRNICVAP30N ~ IRSUICTHNK, design = subset(des, CATAGE == 2), test = "wilcoxon")
```

## Weighted Chi-Squared Tests
We see the association between our categorical confounders and suicidal ideation

```{r survey-weighted chi-squared test between gender and suicidal ideation}
# Test association between two categorical vars
svychisq(~ IRSEX + IRSUICTHNK, design = subset(des,CATAGE == 2), statistic = "adjWald")
# Produce contingency table
prop.table(svytable(~ IRSEX + IRSUICTHNK, subset(des,CATAGE == 2)), margin=2) # col % (weighted)
```

```{r survey-weighted chi-squared test between race and suicidal ideation}
# Test association between two categorical vars
svychisq(~ NEWRACE2 + IRSUICTHNK, design = subset(des,CATAGE == 2), statistic = "adjWald")
# Produce contingency table
prop.table(svytable(~ NEWRACE2 + IRSUICTHNK, subset(des,CATAGE == 2)), margin=2) # col % (weighted)
```

```{r survey-weighted chi-squared test between education and suicidal ideation}
# Test association between two categorical vars
svychisq(~ EDUHIGHCAT + IRSUICTHNK, design = subset(des,CATAGE == 2), statistic = "adjWald")
# Produce contingency table
prop.table(svytable(~ EDUHIGHCAT + IRSUICTHNK, subset(des,CATAGE == 2)), margin=2) # col % (weighted)
```

```{r survey-weighted chi-squared test between IRWRKSTAT18 and suicidal ideation}
# Test association between two categorical vars
svychisq(~ IRWRKSTAT18 + IRSUICTHNK, design = subset(des,CATAGE == 2), statistic = "adjWald")
# Produce contingency table
prop.table(svytable(~ IRWRKSTAT18 + IRSUICTHNK, subset(des,CATAGE == 2)), margin=2) # col % (weighted)
```
```{r survey-weighted chi-squared test between IRPRVHLT and suicidal ideation}
# Test association between two categorical vars
svychisq(~ IRPRVHLT + IRSUICTHNK, design = subset(des,CATAGE == 2), statistic = "adjWald")
# Produce contingency table
prop.table(svytable(~ IRPRVHLT + IRSUICTHNK, subset(des,CATAGE == 2)), margin=2) # col % (weighted)
```

```{r survey-weighted chi-squared test between IRHHSIZ2 and suicidal ideation}
# Test association between two categorical vars
svychisq(~ IRHHSIZ2 + IRSUICTHNK, design = subset(des,CATAGE == 2), statistic = "adjWald")
# Produce contingency table
prop.table(svytable(~ IRHHSIZ2 + IRSUICTHNK, subset(des,CATAGE == 2)), margin=2) # col % (weighted)
```

```{r survey-weighted chi-squared test between INCOME and suicidal ideation}
# Test association between two categorical vars
svychisq(~ INCOME + IRSUICTHNK, design = subset(des,CATAGE == 2), statistic = "adjWald")
# Produce contingency table
prop.table(svytable(~ INCOME + IRSUICTHNK, subset(des,CATAGE == 2)), margin=2) # col % (weighted)
```

```{r survey-weighted chi-squared test between IRDSTNRV12 and suicidal ideation}
# Test association between two categorical vars
svychisq(~ IRDSTNRV12 + IRSUICTHNK, design = subset(des,CATAGE == 2), statistic = "adjWald")
# Produce contingency table
prop.table(svytable(~ IRDSTNRV12 + IRSUICTHNK, subset(des,CATAGE == 2)), margin=2) # col % (weighted)
```

```{r survey-weighted chi-squared test between IRDSTEFF12 and suicidal ideation}
# Test association between two categorical vars
svychisq(~ IRDSTEFF12 + IRSUICTHNK, design = subset(des,CATAGE == 2), statistic = "adjWald")
# Produce contingency table
prop.table(svytable(~ IRDSTEFF12 + IRSUICTHNK, subset(des,CATAGE == 2)), margin=2) # col % (weighted)
```

```{r survey-weighted chi-squared test between IRIMPCONCN and suicidal ideation}
# Test association between two categorical vars
svychisq(~ IRIMPCONCN + IRSUICTHNK, design = subset(des,CATAGE == 2), statistic = "adjWald")
# Produce contingency table
prop.table(svytable(~ IRIMPCONCN + IRSUICTHNK, subset(des,CATAGE == 2)), margin=2) # col % (weighted)
```

```{r survey-weighted chi-squared test between SUTINPPY and suicidal ideation}
# Test association between two categorical vars
svychisq(~ SUTINPPY + IRSUICTHNK, design = subset(des,CATAGE == 2), statistic = "adjWald")
# Produce contingency table
prop.table(svytable(~ SUTINPPY + IRSUICTHNK, subset(des,CATAGE == 2)), margin=2) # col % (weighted)
```