---
title: "DA 401 Data Exploration Final Part"
author: "John Platt"
date: "2025-09-25"
output: 
  html_document:
    code_folding: hide
---

## Introduction
We perform the final part of our data exploration on the 2021-23 NSDUH data, using the following method: weighted 2 sample t-tests. We do this method to capture initial, significant associations between our predictor illicit substance use variables and our response mental health variables. 

## Weighted 2 Sample t-tests
We accomplish the final part of our data exploration.

```{r specifying our survey design, message=FALSE, warning=FALSE}
#Load in cleaned dataset from data folder of repository
nsduh_2123<-read.csv("C:/Users/John Platt/OneDrive/DA301Project/John-Platt-DA-401-Project/data/NSDUH_2021_2023.csv")
#Install survey package
library(survey) #Library for survey design

options(survey.lonely.psu = "adjust")  #common & documented choice to adjust variance if there are single psus within stratum

#Specify survey design
des <- svydesign(
  id      = ~VEREP, #PSU/cluster id
  strata  = ~VESTR_C, #variance strata
  weights = ~ANALWT2_C3, #analysis weight
  data    = nsduh_2123, #2021-2023 NSDUH dataframe
  nest = TRUE
)
```

```{r weighted 2 sample t-test between marijuana use and MDE for young adults, include=TRUE}
#Compare means within CATAGE==2
tt <- svyttest(IRMJFY ~ IRAMDEYR, design = subset(des, CATAGE == 2))
tt
svyby(
  ~IRMJFY, #outcome variable
  ~IRAMDEYR, #grouping variable (two categories)
  subset(des, CATAGE == 2), #subset population
  svymean, #survey weighted mean
  vartype = c("se", "ci"),   #include SEs and 95% CIs
  na.rm=TRUE
  )
```

```{r weighted 2 sample t-test between cocaine use and MDE for young adults, include=TRUE}
#Compare means within CATAGE==2
tt <- svyttest(IRCOCFY ~ IRAMDEYR, design = subset(des, CATAGE == 2))
tt
svyby(
  ~IRCOCFY, #outcome variable
  ~IRAMDEYR, #grouping variable (two categories)
  subset(des, CATAGE == 2), #subset population
  svymean, #survey weighted mean
  vartype = c("se", "ci"),   #include SEs and 95% CIs
  na.rm=TRUE
  )
```

```{r weighted 2 sample t-test between hallucinogen use and MDE for young adults, include=TRUE}
#Compare means within CATAGE==2
tt <- svyttest(IRHALLUCYFQ ~ IRAMDEYR, design = subset(des, CATAGE == 2))
tt
svyby(
  ~IRHALLUCYFQ, #outcome variable
  ~IRAMDEYR, #grouping variable (two categories)
  subset(des, CATAGE == 2), #subset population
  svymean, #survey weighted mean
  vartype = c("se", "ci"),   #include SEs and 95% CIs
  na.rm=TRUE
  )
```

```{r weighted 2 sample t-test between marijuana use and suicidal thoughts for young adults, include=TRUE}
#Compare means within CATAGE==2
tt <- svyttest(IRMJFY ~ IRSUICTHNK, design = subset(des, CATAGE == 2))
tt
svyby(
  ~IRMJFY, #outcome variable
  ~IRSUICTHNK, #grouping variable (two categories)
  subset(des, CATAGE == 2), #subset population
  svymean, #survey weighted mean
  vartype = c("se", "ci"),   #include SEs and 95% CIs
  na.rm=TRUE
  )
#Extract overall standard deviation of IRMJFY
svy_var  <- svyvar(~IRMJFY, design = subset(des, CATAGE == 2), na.rm = TRUE)
#Extract standard deviation
svy_sd <- sqrt(as.numeric(svy_var))
svy_sd
#Extract standard deviation by group
svy_byvar <- svyby(~IRMJFY, ~IRSUICTHNK, design = subset(des, CATAGE == 2), FUN = svyvar, na.rm = TRUE)
svy_byvar$sd <- sqrt(svy_byvar$IRMJFY)
svy_byvar
```

```{r weighted 2 sample t-test between cocaine use and suicidal thoughts for young adults, include=TRUE}
#Compare means within CATAGE==2
tt <- svyttest(IRCOCFY ~ IRSUICTHNK, design = subset(des, CATAGE == 2))
tt
svyby(
  ~IRCOCFY, #outcome variable
  ~IRSUICTHNK, #grouping variable (two categories)
  subset(des, CATAGE == 2), #subset population
  svymean, #survey weighted mean
  vartype = c("se", "ci"),   #include SEs and 95% CIs
  na.rm=TRUE
  )
```

```{r weighted 2 sample t-test between hallucinogen use and suicidal thoughts for young adults, include=TRUE}
#Compare means within CATAGE==2
tt <- svyttest(IRHALLUCYFQ ~ IRSUICTHNK, design = subset(des, CATAGE == 2))
tt
svyby(
  ~IRHALLUCYFQ, #outcome variable
  ~IRSUICTHNK, #grouping variable (two categories)
  subset(des, CATAGE == 2), #subset population
  svymean, #survey weighted mean
  vartype = c("se", "ci"),   #include SEs and 95% CIs
  na.rm=TRUE
  )
```

```{r weighted 2 sample t-test between marijuana use and receiving inpatient mental health treatment for young adults, include=TRUE}
#Compare means within CATAGE==2
tt <- svyttest(IRMJFY ~ MHTINPPY, design = subset(des, CATAGE == 2))
tt
svyby(
  ~IRMJFY, #outcome variable
  ~MHTINPPY, #grouping variable (two categories)
  subset(des, CATAGE == 2), #subset population
  svymean, #survey weighted mean
  vartype = c("se", "ci"),   #include SEs and 95% CIs
  na.rm=TRUE
  )
```

```{r weighted 2 sample t-test between cocaine use and receiving inpatient mental health treatment for young adults, include=TRUE}
#Compare means within CATAGE==2
tt <- svyttest(IRCOCFY ~ MHTINPPY, design = subset(des, CATAGE == 2))
tt
svyby(
  ~IRCOCFY, #outcome variable
  ~MHTINPPY, #grouping variable (two categories)
  subset(des, CATAGE == 2), #subset population
  svymean, #survey weighted mean
  vartype = c("se", "ci"),   #include SEs and 95% CIs
  na.rm=TRUE
  )
```

```{r weighted 2 sample t-test between hallucinogen use and receiving inpatient mental health treatment for young adults, include=TRUE}
#Compare means within CATAGE==2
tt <- svyttest(IRHALLUCYFQ ~ MHTINPPY, design = subset(des, CATAGE == 2))
tt
svyby(
  ~IRHALLUCYFQ, #outcome variable
  ~MHTINPPY, #grouping variable (two categories)
  subset(des, CATAGE == 2), #subset population
  svymean, #survey weighted mean
  vartype = c("se", "ci"),   #include SEs and 95% CIs
  na.rm=TRUE
  )
```

# Weighted Mann-Whitney U Test
Finally, we do a Mann-Whitney U test (non-parametric method) to address our marijuana use distribution not being normal.
```{r Mann-Whitney U Test}
##Survey-weighted Mann–Whitney (Wilcoxon rank-sum) test
svyranktest(IRMJFY ~ IRSUICTHNK, design = subset(des, CATAGE == 2), test = "wilcoxon")
#Table with medians
svyby(~IRMJFY, ~IRSUICTHNK, subset(des, CATAGE == 2), svyquantile, quantiles = 0.5, ci = TRUE)
```