---
title: "Logistic Regression and RCS Analysis"
author: "John Platt"
date: "2025-10-14"
output: 
  html_document:
    code_folding: hide
---

## Introduction
Ensuring assumptions are met beforehand and validating model fit, we conduct survey-weighted logistic regression to examine the association between our main illicit substance use predictors (marijuana use, cocaine use, hallucinogen use) and suicidal ideation. An unadjusted model without confounders will be conducted first, and subsequently, a fully adjusted model will be done.

## Meeting Logistic Regression Assumptions

First, we get our dataset ready and check assumptions before performing logistic regression.

```{r loading in and modifying original dataset, include=TRUE, message=FALSE, warning=FALSE}
#File -> New Project -> Set existing GitHub Repository as working directory with data2 as a subfolder outside the repository and the NSDUH_2023.Rdata file in the subfolder
#Load in the data
load("C:/Users/John Platt/OneDrive/data2/NSDUH_2021_2023.Rdata")
#Required packages:
#dplyr
library(dplyr) #data wrangling and manipulation
#Select a subset of the 2021-2023 NSDUH data that has the 27 necessary columns
data<-data %>% select(ANALWT2_C3,YEAR,VESTR_C,VEREP,IRSEX,CATAGE,NEWRACE2,EDUHIGHCAT,IRWRKSTAT18,IRHHSIZ2,IRPRVHLT,INCOME,IRALCFY,IRMJFY,IRCOCFY,IRCIGFM,IRNICVAP30N,IRHALLUCYFQ,IRALCBNG30D,SUTINPPY,IRDSTNRV12,IRDSTEFF12,IRIMPCONCN,IRSUICTHNK,IRAMDEYR,MHTINPPY,AMISUD5ANYO)
data
# Replace all 91s with 0 in the entire data frame
data[data == 91] <- 0
# Replace all 93s with 0 in the entire data frame
data[data == 93] <- 0
# Replace all 991s with 0 in the entire data frame
data[data == 991] <- 0
# Replace all 993s with 0 in the entire data frame
data[data == 993] <- 0
```

Subsequently, we perform dummy variable coding (for nominal categorical variables) and convert our illicit substance use predictors into categorical variables.

```{r dummy variables and illicit substance use conversion}
#Required packages:
#fastDummies
library(fastDummies) #Dummy variable creation
#Convert nominal variables into dummy variables and drop reference dummy to avoid "dummy variable trap"
data <- dummy_cols(data, select_columns = c("NEWRACE2","IRWRKSTAT18"), remove_first_dummy = TRUE, remove_selected_columns = TRUE)
head(data)
#Convert illicit substance use variables into categories
data$IRMJFY2 <- cut(data$IRMJFY,
                    breaks = 4,   # number of categories
                    labels = c("0-90", "90-180", "180-270", "270-360"),
                    include.lowest = TRUE)
data$IRCOCFY2 <- cut(data$IRCOCFY,
                    breaks = 4,   # number of categories
                    labels = c("0-90", "90-180", "180-270", "270-360"),
                    include.lowest = TRUE)
data$IRHALLUCYFQ2 <- cut(data$IRHALLUCYFQ,
                    breaks = 4,   # number of categories
                    labels = c("0-90", "90-180", "180-270", "270-360"),
                    include.lowest = TRUE)
#Convert categories to numerical codes
data$IRMJFY2 <- ifelse(data$IRMJFY2=="0-90",1,ifelse(data$IRMJFY2=="90-180",2,ifelse(data$IRMJFY2=="180-270",3,ifelse(data$IRMJFY=="270-360",4,NA))))
data$IRCOCFY2 <- ifelse(data$IRCOCFY2=="0-90",1,ifelse(data$IRCOCFY2=="90-180",2,ifelse(data$IRCOCFY2=="180-270",3,ifelse(data$IRCOCFY2=="270-360",4,NA))))
data$IRHALLUCYFQ2 <- ifelse(data$IRHALLUCYFQ2=="0-90",1,ifelse(data$IRHALLUCYFQ2=="90-180",2,ifelse(data$IRHALLUCYFQ2=="180-270",3,ifelse(data$IRHALLUCYFQ2=="270-360",4,NA))))
```

Now we specify our survey design and check strata, PSUs, and effective sample size for our subpopulation.

```{r specify survey design and check sample size as well as strata and PSUs}
#Install survey package
library(survey) #Library for survey design
options(survey.lonely.psu = "adjust")  #common & documented choice to adjust variance if there are single psus within stratum
#Specify survey design
des <- svydesign(
  id      = ~VEREP, #PSU/cluster id
  strata  = ~VESTR_C, #variance strata
  weights = ~ANALWT2_C3, #analysis weight
  data    = data, #2021-2023 NSDUH dataframe
  nest = TRUE
)
#Check strata, PSUs, and effective sample size
young_adults <- subset(des, CATAGE == 2)  # CATAGE==2 for 18–25
# Number of unique strata and PSUs in the subpopulation
length(unique(young_adults$variables$VESTR_C))
length(unique(young_adults$variables$VEREP))
#Compute effective sample size
w <- weights(young_adults)
ess <- (sum(w)^2) / sum(w^2)
ess
```

